Using Metric Space Indexing for Complete and Efficient Record Linkage

### 2017-11-21 paper.tex
Update on Overleaf.

This paper focuses on approaches that use a similarity threshold to
classify links. Such techniques are fundamentally limited by the extent
to which true matching records are similar, and true non-matches are
dissimilar---this is dataset dependent. Within this domain, we define
a technique to be \emph{complete} if it is guaranteed to identify all
pairs within the specified similarity threshold. Commonly used indexing
techniques are incomplete, since they reduce computational cost at the
expense of potentially overlooking some true matches. By definition,
incomplete techniques yield lower recall than complete techniques.
Conversely, and perhaps counter-intuitively, complete techniques can
yield lower precision with some datasets. This is discussed further in
Sect.~\ref{sec-approach}.

@@

A different approach to indexing is the sorted neighborhood
method~\cite{Mon96}, where the datasets to be linked are sorted
according to a \emph{sorting key} (usually a concatenation of the values
from several attributes). A sliding window is then moved over the
datasets and only records within the window are compared. Techniques
that adaptively shrink or expand the window size based on the
characteristics of the sorting key values have shown to improve both
linkage efficiency and quality~\cite{Dra12} over other sliding window
approaches.

@@

Locality sensitive hashing (LSH), originally proposed to allow efficient
nearest-neighbor search in high-dimensional spaces~\cite{Ind98}, has
been employed in record linkage as an indexing technique where attribute
values are hashed multiple times, and blocks are created from those
records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
record linkage approach based on MinHash~\cite{Broder1997} and LSH which
blocks, compares, and then merges linked records in an iterative
fashion, where merged records are re-hashed to improve overall linkage
quality. ~\cite{Steorts2014} evaluates two LSH variations, concluding
that in order to get good results, LSH methods must be tuned to the
particular datasets being linked. This requires good quality ground
truth data which may be unavailable or expensive to obtain.

@@

We address the following general record linkage problem: for two
datasets $A$ and $B$, we wish to find, for each record in $A$, all the
records in $B$ that match it. We compare several algorithms for record
linkage: traditional blocking, one incomplete similarity search method,
LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
also use a simple brute-force complete technique as a baseline for
comparison, although this can only feasibly be applied to the smallest
of our datasets. All experiments have a number of parameters to
configure the search space and algorithm behaviour, including a
\emph{distance threshold} specifying a maximum distance, equivalent to a
minimum similarity, for two records to be classified as a link (i.e.\
referring to the same entity).

@@

\textbf{Traditional blocking:}
The additional parameters are the set of blocking keys and the
normalization encodings applied to each field. These are selected as
described in \cite{Chr12b}, exploiting knowledge of the domain and of
the records being linked, and chosen with the intention of giving the
best possible results for the datasets.
Each record in $A$ is placed into the appropriate block based on its
blocking key value. The algorithm then iterates over the records in $B$,
and for each one, compares it with each of the records from $A$ in the
block with the same blocking key value.

@@

\textbf{THIS TEXT WAS LIFTED FROM BELOW USE HERE? Traditionally, blocking is a method for choosing candidates for linkage using manually designed criteria.
Thanks to blocking, a distance is only computed between promising pairs of records.
A major disadvantage of traditional blocking is the need for extra engineering effort.
Bespoke blocking strategies need to be designed for a given data set and application domain.}

@@

\begin{table}[t]
\caption{Parameter settings used for the different data sets used in
   the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
   shingle size, and the size and number of bands, respectively.}
 \label{table-parameters}
  \centering
  \begin{scriptsize}
  %\addtolength{\tabcolsep}{-0.5pt}
  \begin{tabular}{ccccc}
  \hline\noalign{\smallskip}
  Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
  name(s) & & & &  \\
  \noalign{\smallskip} \hline \noalign{\smallskip}
  CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
    ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
  Isle of Skye &  \\
  Kilmarnock  &  \\
  \noalign{\smallskip} \hline
  \end{tabular}
  \end{scriptsize}
\end{table}

@@

\Cref{comparison-of-results-cora} shows the results of the Cora experiments. The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.  The linker labels may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the size of the bands followed by the number of bands. For traditional blocking the parameters are the fields used for blocking.

TODO: Explain pairs-quality and pairs-completeness.
\emph{\textbf{TODO: discuss measures used. PC, PQ, num comparisons, time? etc.}}

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline{=#=#=}
Brute Force        & 1,677,025                    & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\ \hline
MTree              &  902,693                     & 0.89                  & 0.22                       & 0.84      & 0.57   & 0.68       \\ \hline
LSH-2-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-2           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-2-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-5           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-2-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-10          &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\  \hline
Blocking-year      &  115,893                     & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
Blocking-authors   &   11,039                      & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
Blocking-title     &   27,407                      & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
Blocking-venue     &   36,647                      & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
Blocking-location  & 1,009,957                    & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
Blocking-publisher &  833,079                     & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
Blocking-combined  & 1,214,269                    & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67       \\ \hline
\end{tabular}
\caption{Linkage Quality on the Cora data set distance=70\label{comparison-of-results-cora}
\textbf{TODO: The table needs some rows removed. Note that this table is with a threshold value of 70.}
}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hhline{=#=#=}
MTree    & 102,318,525         & 0.84                  & 0.86                       & 0.65      & 0.46   & 0.54      \\ \hline
LSH-2-2  & 3,109,250           & 0                     & 0.05                       & 0.63      & 0.03   & 0.06       \\
LSH-5-2  & 10,412,496          & 0                     & 0.16                       & 0.64      & 0.11   & 0.19       \\
LSH-10-2 & 53,874,127          & 0                     & 0.46                       & 0.68      & 0.36   & 0.47       \\
LSH-5-5  & 36,566             & 0                     & 0.01                       & 0.76      & 0.01   & 0.01       \\
LSH-10-5 & 129,873            & 0                     & 0.01                       & 0.72      & 0.01   & 0.02       \\ \hline
\end{tabular}
\caption{Linkage Quality on demography datasets \textbf{This is for Skye. Threshold value is 2.}
\label{comparison-of-results-demography-skye}
}
\end{table}

\begin{table}[h]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
MTree     & 514,871,153         & 0.89                  & 0.82                       & 0.76      & 0.45   & 0.57      \\ \hline
LSH-2-2   & 99,145,887          & 0                     & 0.23                       & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338         & 0                     & 0.32                       & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848         & 0                     & 0.48                       & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368            & 0.01                  & 0.02                       & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431            & 0.01                  & 0.03                       & 0.87      & 0.02   & 0.03 \\ \hline
\end{tabular}
\caption{Linkage Quality on demography datasets\textbf{This is for Kilmarnock. Threshold value is 2.
\label{comparison-of-results-demography-kili}
}
\end{table}

>>>>

This paper focuses on approaches that use a similarity threshold to
classify links. Such techniques are fundamentally limited by the extent
to which true matching records are similar, and true non-matches are
dissimilar---this is dataset-dependent. Within this domain, we define
a technique to be \emph{complete} if it is guaranteed to identify all
pairs within the specified similarity threshold. Commonly used indexing
techniques are incomplete, since they reduce computational cost at the
expense of potentially overlooking some true matches. By definition,
incomplete techniques yield lower recall than complete techniques.
Conversely, and perhaps counter-intuitively, complete techniques can
yield lower precision with some datasets. This is discussed further in
Sect.~\ref{sec-approach}.

@@

A different approach to indexing is the sorted neighbourhood
method~\cite{Mon96}, where the datasets to be linked are sorted
according to a \emph{sorting key} (usually a concatenation of the values
from several attributes). A sliding window is then moved over the
datasets and only records within the window are compared. Techniques
that adaptively shrink or expand the window size based on the
characteristics of the sorting key values have shown to improve both
linkage efficiency and quality~\cite{Dra12} over other sliding window
approaches.

@@

Locality sensitive hashing (LSH), originally proposed to allow efficient
nearest-neighbour search in high-dimensional spaces~\cite{Ind98}, has
been employed in record linkage as an indexing technique where attribute
values are hashed multiple times, and blocks are created from those
records that share some hash values. \emph{HARRA}~\cite{Kim10} is a
record linkage approach based on MinHash~\cite{Broder1997} and LSH which
blocks, compares, and then merges linked records in an iterative
fashion, where merged records are re-hashed to improve overall linkage
quality. ~\cite{Steorts2014} evaluates two LSH variations, concluding
that in order to get good results, LSH methods must be tuned to the
particular datasets being linked. This requires good quality ground
truth data which may be unavailable or expensive to obtain.

@@

We address the following general record linkage problem: for two
datasets $A$ and $B$, we wish to find, for each record in $A$, all the
records in $B$ that match it. We compare several algorithms for record
linkage: traditional blocking, one incomplete similarity search method,
LSH-MinHash, and one complete method, M-tree~\cite{paolociaccia2m}. We
also use a simple complete brute-force technique as a baseline for
comparison, although this can only feasibly be applied to the smallest
of our datasets. All experiments have a number of parameters to
configure the search space and algorithm behaviour, including a
\emph{distance threshold} specifying a maximum distance, equivalent to a
minimum similarity, for two records to be classified as a link (i.e.\
referring to the same entity).

@@

\textbf{Traditional blocking:}
The additional parameters are the set of blocking keys and the
normalization encodings applied to each field. These are selected as
described in \cite{Chr12b}, exploiting knowledge of the domain and of
the records being linked, and chosen with the intention of giving the
best possible results for the datasets.
Each record in $A$ is placed into the appropriate block based on its
blocking key value. The algorithm then iterates over the records in $B$,
and for each one, compares it with each of the records from $A$ in the
block with the same blocking key value.
In some circumstances, incomplete approaches such as blocking can yield
higher precision than complete techniques. This can occur when the
datasets contain a significant number
of non-matches that nonetheless have high similarity. In this situation,
the fact that an incomplete technique omits consideration of some
potential links can serve to improve precision, since the similarity
distance

@@

@@

% \begin{table}[t]
% \caption{Parameter settings used for the different data sets used in
%    the experiments,with $d$ the distance threshold used, $l_{ss}$, $l_{sb}$ and $l_{nb}$ the LSH
%    shingle size, and the size and number of bands, respectively.}
%  \label{table-parameters}
%   \centering
%   \begin{scriptsize}
%   %\addtolength{\tabcolsep}{-0.5pt}
%   \begin{tabular}{ccccc}
%   \hline\noalign{\smallskip}
%   Data set~ & $d$ & $l_{ss}$ & $l_{bs}$ & $l_{nb}$ \\
%   name(s) & & & &  \\
%   \noalign{\smallskip} \hline \noalign{\smallskip}
%   CORA & ~$[0,5,10,\ldots,95,100]$~ & ~$2$~ &
%     ~$[2,5,10]$~ & ~$[2,5,10]$~  \\
%   Isle of Skye &  \\
%   Kilmarnock  &  \\
%   \noalign{\smallskip} \hline
%   \end{tabular}
%   \end{scriptsize}
% \end{table}

@@

\Cref{comparison-of-results-cora} shows the results of the Cora experiments. The table shows the parameters for the experiment, the number of distance comparisons made, the precision, recall and F1 measure and the pairs quality and pair completeness achieved by each algorithm.  The linker labels may be interpreted as follows: the algorithm name is followed by its parameters, for LSH this is the number of the bands followed by the size of bands. For traditional blocking the parameters are the fields used for blocking.

TODO: Explain pairs-quality and pairs-completeness.
\emph{\textbf{TODO: discuss measures used. num comparisons, time? etc.}}

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\
\hline
Brute Force        & 1,677,025                    & 0.02                  & 0.33                       & 0.84      & 0.57   & 0.68       \\ \hline
MTree              &  902,693                     & 0.89                  & 0.22                       & 0.84      & 0.57   & 0.68       \\ \hline
LSH-2-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-2            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-2           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-2-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-5            &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-5           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-2-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-5-10           &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\
LSH-10-10          &  175,435                     & 0.31                  & 0.29                       & 0.94      & 0.53   & 0.68       \\  \hline
Blocking-year      &  115,893                     & 0.22                  & 0.19                       & 0.99      & 0.35   & 0.51       \\
Blocking-authors   &   11,039                      & 0.74                  & 0.1                        & 0.94      & 0.16   & 0.28       \\
Blocking-title     &   27,407                      & 0.84                  & 0.23                       & 0.95      & 0.42   & 0.58       \\
Blocking-venue     &   36,647                      & 0.7                   & 0.12                       & 0.85      & 0.29   & 0.44       \\
Blocking-location  & 1,009,957                    & 0.17                  & 0.24                       & 0.83      & 0.43   & 0.57       \\
Blocking-publisher &  833,079                     & 0.12                  & 0.23                       & 0.85      & 0.44   & 0.58       \\
Blocking-combined  & 1,214,269                    & 0.05                  & 0.32                       & 0.84      & 0.56   & 0.67       \\ \hline
\end{tabular}
\caption{Linkage Quality on the Cora data set distance=70\label{comparison-of-results-cora}
\textbf{TODO: The table needs some rows removed. Note that this table is with a threshold value of 70.}
}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
MTree    & 102,318,525         & 0.84                  & 0.86                       & 0.65      & 0.46   & 0.54      \\ \hline
LSH-2-2  & 3,109,250           & 0                     & 0.05                       & 0.63      & 0.03   & 0.06       \\
LSH-5-2  & 10,412,496          & 0                     & 0.16                       & 0.64      & 0.11   & 0.19       \\
LSH-10-2 & 53,874,127          & 0                     & 0.46                       & 0.68      & 0.36   & 0.47       \\
LSH-5-5  & 36,566             & 0                     & 0.01                       & 0.76      & 0.01   & 0.01       \\
LSH-10-5 & 129,873            & 0                     & 0.01                       & 0.72      & 0.01   & 0.02       \\ \hline
\end{tabular}
\caption{Linkage Quality on demography datasets \textbf{This is for Skye. Threshold value is 2.}
\label{comparison-of-results-demography-skye}
}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{l|r|r|r|r|r|r}
Linker   & Comparisons & Pairs quality & Pairs completeness & Precision & Recall & F1 Measure \\ \hline
MTree     & 514,871,153         & 0.89                  & 0.82                       & 0.76      & 0.45   & 0.57      \\ \hline
LSH-2-2   & 99,145,887          & 0                     & 0.23                       & 0.81      & 0.16   & 0.27       \\
LSH-5-2   & 130,721,338         & 0                     & 0.32                       & 0.79      & 0.23   & 0.36       \\
LSH-10-2  & 177,168,848         & 0                     & 0.48                       & 0.79      & 0.36   & 0.49       \\
LSH-5-5   & 239,368            & 0.01                  & 0.02                       & 0.84      & 0.01   & 0.02       \\
LSH-10-5  & 855,431            & 0.01                  & 0.03                       & 0.87      & 0.02   & 0.03 \\ \hline
\end{tabular}
\caption{Linkage Quality on demography datasets\textbf{This is for Kilmarnock. Threshold value is 2.}
\label{comparison-of-results-demography-kili}
}
\end{table}
